---
title: '**overdispsim** - Simulations of Overdispersion in SECR'
author: "Murray Efford and David Fletcher"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
    toc_depth: 2
  pdf_document: 
    toc: yes
    toc_depth: 2
vignette: > 
  %\VignetteIndexEntry{Simulations of Overdispersion in SECR} 
  %\VignetteEngine{knitr::rmarkdown} 
  \usepackage[utf8]{inputenc} 
---

# Introduction

This vignette explains some of the functions in the package **overdispsim** and 
provides code for simulations in the paper of Efford and Fletcher (2024). 

Simulations are re-run if the variable 'runsimulations' is set to TRUE; 
otherwise the simulations are downloaded from Zenodo (Efford 2025).

**overdispsim** uses the R packages **secr** (Efford 2023a) and 
**secrdesign** (Efford 2023b) available from the CRAN repository. 

Three processes are simulated for the distribution of activity centres (AC): a log-Gaussian Cox process (LGCP), a Thomas cluster process and a random binary habitat mosaic.

Functions `rLGCP` and `rThomas` from the **spatstat** package (Baddeley et al. 2015) are used where possible. The algorithm for `rLGCP` in **spatstat.random** changed in version 3.2.1 (Oct 2023) to avoid dependence on **RandomFields**. Simulations here were run with the earlier code and may differ (see warning in `?spatstat.random::rLGCP`).

However, simulations conditional on $N(A)$ (the fixed number in buffered area $A$) required custom code for LGCP (the exported function `myRF`). Package **RandomFields** (Schlather et al. 2015) is required for LGCP simulations 
with `myRF`. Package **RandomFields** is not currently on CRAN. It may be installed with this code:
	
```
install.packages("RandomFields", repos = "https://spatstat.r-universe.dev")
```

The fixed cluster process of Bischof et al. (2020) was emulated using the **secr** function `sim.popn` with settings model2D = "cluster" and details$clone = "constant" (see below).

Package **zen4R** (Blondel 2024) is required to download simulations from Zenodo.

# Setup

```{r setup, warning = FALSE, message = FALSE}
library(overdispsim, quietly = TRUE)
library(zen4R)
options(digits = 6)
runsimulations <- FALSE   # set TRUE to run afresh
runexamples <- FALSE
```

We next set some options by assigning variables to an environment ('.local') that is 
available to other functions in the package, and incidentally keep a copy 
in a list named 'localpar'. 

The function is run with these arguments automatically when the package is attached 
(i.e. in .onAttach), so these are default settings. The actual number of cores 
is capped at the number available if that is less than the given maxncores.

```{r setup2}

localpar <- setlocalparameters(
	lambda0     = 0.5,
	sigma       = 1.0,
	detectfn    = 'HHN',
	noccasions  = 5,
	traps       = make.grid(12,12, detector = 'proximity', spacing = 2.0),
	maskspacing = 0.5,
	maskbuffer  = 4,
	N           = 256,
	maxncores   = 18
)
```

# Interesting derived quantities

These are non-essential.

```{r derived}

# expected variance of binomial n
a <- sum(localpar$pd * attr(localpar$mask, 'area'))
p <- a / maskarea(localpar$mask)
Evarn <- p * (1-p) * localpar$N

# expected number of individuals detected
En <- sum(localpar$pd) * attr(localpar$mask, 'area') * localpar$D

# expected number of individuals at each detector
enk <- Enk(D = localpar$D, localpar$mask, localpar$traps, detectfn = localpar$detectfn,
		   detectpar = localpar$detectpar, noccasions = localpar$noccasions)
```

# Define populations

Start by specifying some vectors with varying levels of parameters:
```{r levels}
Alevels  <- c(0.0625, 0.125, 0.25, 0.5, 0.75, 1.0) # randomHabitat A
Vlevels  <- c(0, 0.125, 0.25, 0.5, 0.75, 1.0)      # LGCP variances
mulevels <- c(1,2,4,8,16,32)                       # expected number per cluster   
```

## Poisson N(A)

```{r definepop}
# LGCP
basepopargs1 <- list(
	D = localpar$D, 
	core = localpar$mask,
	buffer = 0,    # uses bounding box of mask
	model2D = "rLGCP",
	details = list(var = 0, scale = 5 * localpar$sigma, eps = spacing(localpar$mask), 
				   saveLambda = TRUE))
popargs1 <- extend(basepopargs1, values = list(var = Vlevels, 
								scale = c(5,10) * localpar$sigma))
#-------------------------------------------------------------------------------

# Thomas clustering
basepopargs2 <- list(D = localpar$D, 
					 core = localpar$mask,
					 buffer = spacing(localpar$mask)/2,
					 model2D = "rThomas",
					 details = list(mu = 1, scale = localpar$sigma, eps = 
					 			   	spacing(localpar$mask), saveLambda = TRUE))
# rThomas requires scale>0
popargs2 <- extend(basepopargs2, values = list(mu = mulevels, 
								scale = c(1e-4, 1, 2, 4)*localpar$sigma))

#-------------------------------------------------------------------------------

# random habitat 
basepopargs3 <- list(D = randomDensity, 
					 core = localpar$mask,
					 buffer = 0,
					 model2D = "IHP",
					 details = list(D = localpar$D, p=0.5, A=0.25, rescale=TRUE))
popargs3 <- extend(basepopargs3, values = list(A = Alevels, p = c(0.25, 0.5)) )
```

## Fixed N(A)

```{r definepopfixed}

# random field
basepopargs1f <- list(
	D = myRF, 
	core = localpar$mask,
	buffer = 0,
	model2D = "IHP",
	Ndist   = "fixed",
	Nbuffer = localpar$N,
	details = list(var = 1, scale = 1, N = localpar$N, eps = spacing(localpar$mask), 
				   saveLambda = TRUE))
popargs1f <- extend(basepopargs1f, values = list(var = Vlevels,
								scale = c(5,10) * localpar$sigma))
#-------------------------------------------------------------------------------

# Clusters (constant size)
# no algorithm for fixed-N rThomas?
# cf Bischof et al. fixed clusters (clone = "constant", scale = 0)

basepopargs2f <- list(D = localpar$D, 
					  core = localpar$mask,
					  buffer = spacing(localpar$mask)/2,
					  model2D = "cluster",
					  Ndist   = "fixed",
					  Nbuffer = localpar$N,
					  details = list(mu = 1, scale = 1, clone = "constant",
					  	eps = spacing(localpar$mask), saveLambda = TRUE))
popargs2f <- extend(basepopargs2f, values = list(mu = mulevels, 
								scale = c(0,1,2,4)*localpar$sigma))
#-------------------------------------------------------------------------------

# random habitat, fixed-N
basepopargs3f <- list(D      = randomDensity, 
					  core    = localpar$mask,
					  buffer  = 0,
					  model2D = "IHP",
					  Ndist   = "fixed",
					  Nbuffer = localpar$N,
					  details = list(D = D, p=0.5, A=0.25, rescale=TRUE))
popargs3f <- extend(basepopargs3f, values = list(A = Alevels, p = c(0.25, 0.5)) )
```

# Example

Select a subset of cluster (Thomas process) population scenarios and extract the parameter levels for later use. 
```{r pops}
sapply(popargs2, '[[', 'details')[1:2,]  # display parameter values
pops <- popargs2[13:18]                  # select mu = 1..32, scale = 2
```

And plot an example with mu = 32 and scale = 2:
```{r pop plot}
set.seed(12345)
pop <- do.call(sim.popn, popargs2[[18]])
plot(pop)                   # AC
plot(localpar$traps, add = TRUE)   # detectors
```

Further execution of the example is suppressed by default, but can be cut-and-pasted into an R session. 

## No model fit

Messages are suppressed.

```{r example1, message = FALSE, cache = TRUE, eval = runexamples}
sims <- run_all(nrepl = 100, pops, fit = FALSE)
sumn <- summary_n(sims)
sumn[,1:10]
```

## Model fit

Using the same population scenarios -
```{r example2, message = FALSE, cache = TRUE, warning = FALSE, eval = runexamples}
simsM <- run_all(nrepl = 10, pops, fit = TRUE)  
summary_M(simsM)                     
```

# Code for simulations in paper

Here we list the commands used to generate the main results. Each of these takes substantial time to run (hours). They should be run separately and saved for later processing (i.e. summary).

| Code | AC distribution | Fit | extractfn | Note |
-----|-----------|-----------|-----------|-----------------
1 | LGCP | none | extract_n |
2 | Thomas process | none |extract_n |
3 | random habitat | none |extract_n |
1f | fixed-N(A) LGCP | none |extract_n |
2f | fixed cluster process | none |extract_n |
3f | fixed-N(A) random habitat | none |extract_n |
1M | LGCP | full likelihood | extract_M |
2M | Thomas process | full likelihood |  extract_M |
3M | random habitat | full likelihood | extract_M |
1Mf | fixed-N(A) LGCP | full likelihood | extract_M |
2Mf | fixed cluster process | full likelihood | extract_M |
3Mf | fixed-N(A) random habitat | full likelihood | extract_M |
1MCL | LGCP | conditional likelihood | extract_MCL |
2MCL | Thomas process | conditional likelihood |  extract_MCL |
3MCL | random habitat | conditional likelihood | extract_MCL |

## Sampling only

Here we generate SECR samples, count the number of detected individuals $n$ and 
compute $\hat c$ using the known detection parameters. Models are not fitted.

```{r samplingonly, eval = runsimulations}

nrepl <- 10000

# Poisson N(A)
sims1 <- run_all(nrepl, popargs1, fit = FALSE, byscenario = FALSE)
sims2 <- run_all(nrepl, popargs2, fit = FALSE, byscenario = TRUE)
sims3 <- run_all(nrepl, popargs3, fit = FALSE, byscenario = TRUE)

# Fixed N(A)
sims1f <- run_all(nrepl, popargs1f, fit = FALSE, byscenario = FALSE)
# byscenarios = TRUE throws error
sims2f <- run_all(nrepl, popargs2f, fit = FALSE, byscenario = TRUE)
sims3f <- run_all(nrepl, popargs3f, fit = FALSE, byscenario = TRUE)
```

## Full model fits (suffix M)

```{r modelfits, eval = runsimulations}

nreplM <- 1000

# Poisson N(A)
sims1M <- run_all(nreplM, popargs1[7:12],  fit = TRUE)
sims2M <- run_all(nreplM, popargs2[13:18], fit = TRUE, 
			  start = list(D = 3000, lambda0 = 0.4, sigma = 2.2))
sims3M <- run_all(nreplM, popargs3[7:12], fit = TRUE)

# Fixed N(A)
sims1fM <- run_all(nreplM, popargs1f[7:12], fit = TRUE, distribution = "binomial", 
			   start = list(D = 5000))
sims3fM <- run_all(nreplM, popargs3f[7:12], fit = TRUE, distribution = "binomial")
sims2fM <- run_all(nreplM, popargs2f[7:12], fit = TRUE, distribution = "binomial")
```

## Summaries

Simulation results have been archived on Zenodo (Efford 2025). If necessary, 
we retrieve them with the R package **zen4R** (Blondel 2024).

```{r zenodo, eval = !runsimulations, echo = TRUE}
tmpfolder <- tempdir()
zen4R::download_zenodo("10.5281/zenodo.14873670", path = tmpfolder, quiet = TRUE)
files <- list.files(path = tmpfolder, full.names = TRUE)
files <- files[grepl('.rdata', tolower(files))]
reloaded <- unlist(lapply(files, load, envir = .GlobalEnv))
# objects successfully reloaded:
print(reloaded)
```

Now summarise the simulations

```{r summaryM, eval = TRUE}

# no fit
sum1  <- summary_n(sims1)
sum2  <- summary_n(sims2)
sum3  <- summary_n(sims3)
sum1f  <- summary_n(sims1f)
sum2f  <- summary_n(sims2f)
sum3f  <- summary_n(sims3f)

# vs detection-weighted local density
sum1M  <- summary_M(sims1M)
sum2M  <- summary_M(sims2M)
sum3M  <- summary_M(sims3M)

# vs global density
sum1MT <- summary_M(sims1M, true = localpar$D)
sum2MT <- summary_M(sims2M, true = localpar$D)
sum3MT <- summary_M(sims3M, true = localpar$D)

# conditional likelihood (for COV(a) etc.)
sum1MCL <- summary_MCL(sims1MCL, true = localpar$D)
sum2MCL <- summary_MCL(sims2MCL, true = localpar$D)
sum3MCL <- summary_MCL(sims3MCL, true = localpar$D)
```

```{r savesummary, echo = FALSE, eval = FALSE}
# summaries were previously saved to data folder with this code

# no fit
save(sum1, file = 'data/sum1.RData')
save(sum2, file = 'data/sum2.RData')
save(sum3, file = 'data/sum3.RData')

# no fit, fixed
save(sum1f, file = 'data/sum1f.RData')
save(sum2f, file = 'data/sum2f.RData')
save(sum3f, file = 'data/sum3f.RData')

# vs detection-weighted local density
save(sum1M, file = 'data/sum1M.RData')
save(sum2M, file = 'data/sum2M.RData')
save(sum3M, file = 'data/sum3M.RData')

# vs global density
save(sum1MT, file = 'data/sum1MT.RData')
save(sum2MT, file = 'data/sum2MT.RData')
save(sum3MT, file = 'data/sum3MT.RData')

# conditional likelihood (for COV(a) etc.)
save(sum1MCL, file = 'data/sum1MCL.RData')
save(sum2MCL, file = 'data/sum2MCL.RData')
save(sum3MCL, file = 'data/sum3MCL.RData')

```

...and output to .csv.

```{r csv, eval = FALSE, echo = FALSE}
write.csv(sum1M, file='extra/sum1M.csv')
write.csv(sum2M, file='extra/sum2M.csv')
write.csv(sum3M, file='extra/sum3M.csv')

write.csv(sum1MT, file='extra/sum1MT.csv')
write.csv(sum2MT, file='extra/sum2MT.csv')
write.csv(sum3MT, file='extra/sum3MT.csv')

write.csv(sum1MCL, file='extra/sum1MCL.csv')
write.csv(sum2MCL, file='extra/sum2MCL.csv')
write.csv(sum3MCL, file='extra/sum3MCL.csv')
```

# CL

The SECR model was fitted by maximising the conditional likelihood as a convenient
way in **secr** to simulate the coverage of confidence intervals for the effective
sampling area $a(\hat \theta)$ and the proportion of variance due to $n$ and $a(\hat \theta)$.

```{r CLfit, eval = runsimulations}
# Poisson N
sims1MCL <- run_all(1000, popargs1[7:12], fit = TRUE, CL = TRUE)
sims2MCL <- run_all(1000, popargs2[13:18], fit = TRUE, CL = TRUE,
				 start = list(D = 3000, lambda0 = 0.4, sigma = 2.2))
sims3MCL <- run_all(1000, popargs3[7:12], fit = TRUE, CL = TRUE)
```

# Cohesion

Additional simulations were performed with overdispersion in the detection process. A novel function (`sim.capthist`) is defined for simulating detection with variable within-cluster cohesion, ranging from none (gamma = 0) to complete (gamma = 1). This applies *only* to clustered AC.

```{r cohesion, eval = runsimulations}
detargs <- list (savepopn = TRUE, gamma = 1)  # complete cohesion
# No fit
sims2C <- run_all(10000, popargs2, CH.function = "sim.cohesion", detargs = detargs,
			   fit = FALSE, byscenario = TRUE)
sims2Cf <- run_all(10000, popargs2f, CH.function = "sim.cohesion", detargs = detargs,
			   fit = FALSE, byscenario = TRUE)
# Conditional likelihood fit (effect of cohesion on coverage of $\hat a$)
sims2CCL <- run_all(1000, popargs2[1:6], CH.function = "sim.cohesion", detargs = detargs,
			   fit = TRUE, CL = TRUE, byscenario = FALSE)
sims2CfCL <- run_all(1000, popargs2f[1:6], CH.function = "sim.cohesion", detargs = detargs,
				fit = TRUE, CL = TRUE, byscenario = FALSE)
```

```{r cohesion1, eval = FALSE, echo = FALSE}
sum2CCL <- summary_MCL(sims2CCL, true = localpar$D)
sum2CfCL <- summary_MCL(sims2CfCL, true = localpar$D)
```

# References

Baddeley, A., Rubak, E., and Turner, R. (2015) *Spatial Point Patterns: Methodology and Applications with R*. Chapman and Hall/CRC Press, London.

Bischof, R., Dupont, P., Milleret, C., Chipperfield, J., and Royle, J. A. (2020) Consequences of ignoring group association in spatial capture--recapture analysis. *Wildlife Biology* wlb.00649. DOI 10.2981/wlb.00649

Blondel, E. (2024) zen4R: Interface to 'Zenodo' REST API. R package version
  0.10. https://CRAN.R-project.org/package=zen4R/

Borchers, D. L. and Efford, M. G. (2008) Spatially explicit maximum likelihood methods for capture--recapture studies. *Biometrics* **64**, 377--385.

Efford, M. G. (2023a) secr: Spatially explicit capture--recapture models. R package version
  4.6.4. https://CRAN.R-project.org/package=secr/

Efford, M. G. (2023b) openCR: Spatially explicit capture--recapture models. R package version
  2.2.6. https://CRAN.R-project.org/package=openCR/

Efford, M. G. (2025) Simulations of overdispersed activity centres in spatially explicit capture-recapture [Data set]. Zenodo. https://doi.org/10.5281/zenodo.14873670

Efford, M. G. and Fletcher, D. (2024)  The effect of spatial overdispersion on confidence intervals for population density estimated by spatially explicit capture--recapture. bioRxiv https://doi.org/10.1101/2024.03.12.584742.

Saura, S., and Martínez-Millán, J. (2000) Landscape patterns simulation with a modified random clusters method. *Landscape Ecology* **15**, 661--678.

Schlather, M., Malinowski, A., Menck, P. J., Oesting, M., and Strokorb, K. (2015) Analysis, simulation and prediction of multivariate random fields with package RandomFields. *Journal of Statistical Software* **63**, 1--25. https://www.jstatsoft.org/v63/i08/

# Appendix 1. Simulation summaries {#appendix1}

## No model

```{r summarynprint, eval = TRUE}
summary_n(sims1) # LGCP Poisson N(A)
summary_n(sims2) # Thomas Poisson N(A)
summary_n(sims3) # Random habitat Poisson N(A)

summary_n(sims1f) # LGCP fixed N(A)
summary_n(sims2f) # Fixed cluster process
summary_n(sims3f) # Random habitat fixed N(A)
```

## Model fitted

Summarise relative to local density --
```{r summaryMprint1, eval = TRUE}
summary_M(sims1M) # LGCP Poisson N(A)
summary_M(sims2M) # Thomas Poisson N(A)
summary_M(sims3M) # Random habitat Poisson N(A)

summary_M(sims1fM) # LGCP fixed N(A)
summary_M(sims2fM) # Fixed cluster process
summary_M(sims3fM) # Random habitat fixed N(A)
```

Summarise relative to global density --
```{r summaryMprint2, eval = TRUE}
summary_M(sims1M, true = localpar$D) # LGCP Poisson N(A)
summary_M(sims2M, true = localpar$D) # Thomas Poisson N(A)
summary_M(sims3M, true = localpar$D) # Random habitat Poisson N(A)
```

## Model fitted, conditional likelihood

```{r summaryMCLprint, eval = TRUE}
summary_MCL(sims1MCL) # LGCP Poisson N(A)
summary_MCL(sims2MCL) # Thomas Poisson N(A)
summary_MCL(sims3MCL) # Random habitat Poisson N(A)
```

## Model fitted, conditional likelihood, complete cohesion

```{r cohesion2, eval = TRUE}
summary_MCL(sims2CCL, true = localpar$D)   # Thomas cluster process
summary_MCL(sims2CfCL, true = localpar$D)  # fixed cluster process
```

# Finally...

```{r cleanup, eval = !runsimulations}
# appears to fail when knitting to pdf
unlink(tmpfolder, recursive = TRUE)
```

